---
title: "DCA_DataFrames"
author: "Ben J. Rivera"
date: '2023-01-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(here)
library(rtry)


```

The goal of this script is to create and clean the necessary dataframes to be used in analysis later. Run this first then move on to other scripts.


# Reading CSVs in
Okay, this is the boring part. I need to just load in all my datasets and get everything nice and pretty. I'm going to have to deal with plant names.

## Places

Loading in data based on where it was collected

```{r}
Proper_Names <- read_csv(here("EnteredData", "PlantNameChanger.csv"))


# Ten Mile
TeMi_Coords <- read_csv(here("EnteredData", "TeMi_RelativeCoords.csv"))
TeMi_Dist <- read_csv(here("EnteredData", "TeMi_Distance.csv"))
TeMi_Locations <- read_csv(here("EnteredData", "TeMi_Location.csv"))


# Zmudowski
Zmud_Coords <- read_csv(here("EnteredData", "Zmud_RelativeCoords.csv"))
Zmud_Dist <- read_csv(here("EnteredData", "Zmud_Distance.csv"))
Zmud_Locations <- read_csv(here("EnteredData", "ZmudLocations.csv"))


# Guadalupe Nipomo
GuNi_Coords <- read_csv(here("EnteredData", "GuNi_RelativeCoords.csv"))
GuNi_Dist <- read_csv(here("EnteredData", "GuNi_Distance.csv"))
GuNi_Locations <- read_csv(here("EnteredData", "GuNi_Location.csv"))


```

### Fixing Cover Names
```{r}
TeMi_Cover <- read_csv(here("EnteredData", "TeMi_Cover.csv"))
TeMi_Cover<- merge(TeMi_Cover, Proper_Names, by.x = "Species", by.y = "Old_Species_Name")
rownames(TeMi_Cover)<- TeMi_Cover$New_Species_Name
TeMi_Cover<- TeMi_Cover[2:(length(TeMi_Cover)-1) ]

Zmud_Cover <- read_csv(here("EnteredData", "Zmud_Cover.csv"))
Zmud_Cover<- merge(Zmud_Cover, Proper_Names, by.x = "Species", by.y = "Old_Species_Name")
rownames(Zmud_Cover)<- Zmud_Cover$New_Species_Name
Zmud_Cover<- Zmud_Cover[2:(length(Zmud_Cover)-1) ]

GuNi_Cover <- read_csv(here("EnteredData", "GuNi_Cover.csv"))
GuNi_Cover<- merge(GuNi_Cover, Proper_Names, by.x = "Species", by.y = "Old_Species_Name")
rownames(GuNi_Cover)<- GuNi_Cover$New_Species_Name
GuNi_Cover<- GuNi_Cover[2:(length(GuNi_Cover)-1) ]

```


### Create Relative Coords

Here I am creating coordinates for each point at each site relative to the plot. Rather than being absolute, North South, these are positioned parrelel to the waterline. This will be used for drawing environmental comparisons along these axes which make more sense with stress.

```{r}
#TeMi
TeMi_Subset_X <- subset(TeMi_Coords, IN_FID == 1) #Just the X 
TeMi_Subset_Y <- subset(TeMi_Coords, IN_FID == 2) #Just the Y
TeMi_Sub<-merge(TeMi_Subset_X, TeMi_Subset_Y, by = "NEAR_FID") #recombine them
TeMi_RelCords<- data.frame(Plot= TeMi_Sub$NEAR_FID, X = TeMi_Sub$NEAR_DIST.x, Y = TeMi_Sub$NEAR_DIST.y, Z = TeMi_Locations$POINT_Z) #make a data frame with XYZ
TeMi_RelCords$Plot<- merge(TeMi_RelCords[1], TeMi_Locations[1:2], by.x = "Plot", by.y = "OID_")$Comment #This fixes the names
TeMi_RelCords<- mutate(TeMi_RelCords, X = ifelse(TeMi_RelCords$Plot == "WATERLINE",-1*X, X)) #
TeMi_RelCords$Z <- TeMi_RelCords$Z - min(subset(TeMi_RelCords,Plot != "WATERLINE")$Z)

#Zmud
Zmud_Subset_X <- subset(Zmud_Coords, IN_FID == 2)
Zmud_Subset_Y <- subset(Zmud_Coords, IN_FID == 3)
Zmud_Sub<-merge(Zmud_Subset_X, Zmud_Subset_Y, by = "NEAR_FID")
Zmud_RelCords<- data.frame(Plot= Zmud_Sub$NEAR_FID, X = Zmud_Sub$NEAR_DIST.x, Y = Zmud_Sub$NEAR_DIST.y, Z = Zmud_Locations$POINT_Z)
Zmud_RelCords$Plot<- merge(Zmud_RelCords[1], Zmud_Locations[1:2], by.x = "Plot", by.y = "OID_")$Comment
Zmud_RelCords<- mutate(Zmud_RelCords, X = ifelse(Zmud_RelCords$Plot == "WATERLINE",-1*X, X))
Zmud_RelCords$Z <- Zmud_RelCords$Z - min(subset(Zmud_RelCords,Plot != "WATERLINE")$Z)

#GuNi
GuNi_Subset_X <- subset(GuNi_Coords, IN_FID == 3)
GuNi_Subset_Y <- subset(GuNi_Coords, IN_FID == 2)
GuNi_Sub<-merge(GuNi_Subset_X, GuNi_Subset_Y, by = "NEAR_FID")
GuNi_RelCords<- data.frame(Plot= GuNi_Sub$NEAR_FID, X = GuNi_Sub$NEAR_DIST.x, Y = GuNi_Sub$NEAR_DIST.y, Z = GuNi_Locations$POINT_Z)
GuNi_RelCords$Plot<- merge(GuNi_RelCords[1], GuNi_Locations[1:2], by.x = "Plot", by.y = "OID_")$Comment
GuNi_RelCords<- mutate(GuNi_RelCords, X = ifelse(GuNi_RelCords$Plot == "WATERLINE",-1*X, X))
GuNi_RelCords$Z <- GuNi_RelCords$Z - min(GuNi_RelCords$Z)




```


### Checking Coords

Here I am making sure that the rotation worked properly. 

```{r}
ggplot(aes(x = X, y = Y, label = Plot, color = Z), data = TeMi_RelCords)+
  geom_point()+
  geom_label()

ggplot(aes(x = POINT_X, y = POINT_Y, label = Comment, color = POINT_Z), data = TeMi_Locations)+
  geom_point()+
  geom_text()+
  ggtitle("Ten Mile")


ggplot(aes(x = X, y = Y, label = Plot, color = Z), data = Zmud_RelCords)+
  geom_point()+
  geom_label()
ggplot(aes(x = POINT_X, y = POINT_Y, label = Comment, color = POINT_Z), data = Zmud_Locations)+
  geom_point()+
  geom_text()+
  ggtitle("Zmudowski")

ggplot(aes(x = X, y = Y, label = Plot, color = Z), data = GuNi_RelCords)+
  geom_point()+
  geom_label()
ggplot(aes(x = POINT_X, y = POINT_Y, label = Comment, color = POINT_Z), data = GuNi_Locations)+
  geom_point()+
  geom_text()+
  ggtitle("Guadalupe-Nipomo")

```




# Spanning Dataframes

## Env Data


```{r}
DuneEnvData <- read_csv(here("EnteredData", "DuneEnvData.csv"))

ggplot(aes(x= SoilTemp, y = SoilMoist, color = Site), data = DuneEnvData)+
  geom_point() #okay so site super mattered which makes sense for temp. I'll scale it

ggplot(aes(x= scale(SoilTemp), y = SoilMoist, color = Site), data = DuneEnvData)+
  geom_point() #Lol. I need to scale Soil Temp by the site! Duh! I should actually do that for Soil Moisture too but it won't be as severe


DuneEnv_scaled <- DuneEnvData %>% group_by(Site) %>% mutate(SoilTemp = scale(SoilTemp), SoilMoist=scale(SoilMoist))


DuneEnv_scaled<-plyr::ddply(DuneEnvData, "Site", transform, SoilTemp = scale(SoilTemp))
DuneEnv_scaled<-plyr::ddply(DuneEnv_scaled, "Site", transform, SoilMoist = scale(SoilMoist))


#Temp v moist
ggplot(aes(x= SoilTemp, y = SoilMoist), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) +
  geom_smooth(method = "lm", aes(color = Site), fill = NA)+
  geom_smooth(method = "lm", color = "black")

cor.test(x = DuneEnv_scaled$SoilMoist, y = DuneEnv_scaled$SoilTemp, method = "pearson") # temp and mositure are not correlated which is nice

# Slope v SoilMoist
ggplot(aes(x= Slope, y = SoilMoist), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) +
  geom_smooth(method = "lm")# No obvious pattern
cor.test(x = DuneEnv_scaled$Slope, y = DuneEnv_scaled$SoilMoist, method = "pearson") # Nutthin

#Slope v Temp ***
ggplot(aes(x= Slope , y =SoilTemp), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) +
  geom_smooth(method = "lm")
cor.test(x = DuneEnv_scaled$Slope, y = DuneEnv_scaled$SoilTemp, method = "pearson") # Topsoil, Slope explains SoilTemp

#aspect slope
ggplot(aes(x= Aspect, y = Slope), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) # No obvious pattern
summary(aov(Slope ~ Aspect, data = DuneEnv_scaled)) #nuthin


#aspect temp
ggplot(aes(x= Aspect, y = SoilTemp), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) # No obvious pattern
summary(aov(SoilTemp ~ Aspect, data = DuneEnv_scaled)) #nuthin


#aspect Moist
ggplot(aes(x= Aspect, y = SoilMoist), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) # No obvious pattern
summary(aov(SoilMoist ~ Aspect, data = DuneEnv_scaled)) #nuthin
```




## Phenology

```{r}
Pheno <-read_csv(here("EnteredData", "DunePhenoTrait.csv"))

head(Pheno)

CollectedTraits <- subset(Pheno,  is.na(LeafLength) == FALSE)


SuccTraits <- rbind(subset(CollectedTraits, Species == "Yellow Sand Verbena"), subset(CollectedTraits, Species == "Pink Sand Verbena"), subset(CollectedTraits, Species == "Purple Sand Verbena"), subset(CollectedTraits, Species == "Carpobrotus edulis"))

```


### Checking Graphs

```{r}
hist(CollectedTraits$LeafLength) #Normal
hist(CollectedTraits$LeafWidth) #might need Log
hist(log(CollectedTraits$LeafWidth)) #not really better
hist(CollectedTraits$LeafThickness) #typo and log transform
hist(log(CollectedTraits$LeafThickness))
hist(CollectedTraits$Height) #
hist(CollectedTraits$BasalDiam)


subset(CollectedTraits, LeafTraits$LeafThickness >20) #Thats a typo
subset(CollectedTraits, LeafTraits$Height >205) #Okay, maybe legit
subset(CollectedTraits, LeafTraits$BasalDiam >25) #Just a chonky boi but worth double checking


```

```{r}
ggplot(aes(x = LeafLength, y = LeafWidth, color = Species, shape = Site), data = SuccTraits)+
  geom_point()
ggplot(aes(x = LeafLength, y = LeafThickness, color = Species, shape = Site), data = SuccTraits)+
  geom_point()
ggplot(aes(x = LeafLength, y = Height, color = Species, shape = Site), data = SuccTraits)+
  geom_point()
ggplot(aes(x = LeafLength, y = BasalDiam, color = Species, shape = Site), data = SuccTraits)+
  geom_point()

ggplot(aes(x = LeafWidth, y = LeafThickness, color = Species, shape = Site), data = SuccTraits)+
  geom_point()
ggplot(aes(x = LeafWidth, y = Height, color = Species, shape = Site), data = SuccTraits)+
  geom_point()
ggplot(aes(x = LeafWidth, y = BasalDiam, color = Species, shape = Site), data = SuccTraits)+
  geom_point()

ggplot(aes(x = LeafThickness, y = Height, color = Species, shape = Site), data = SuccTraits)+
  geom_point()
ggplot(aes(x = LeafThickness, y = BasalDiam, color = Species, shape = Site), data = SuccTraits)+
  geom_point()

ggplot(aes(x = Height, y = BasalDiam, color = Species, shape = Site), data = SuccTraits)+
  geom_point()


```

# Traits From Databases

Okay, let's figure out how to get some traits from the online world. There is code back from the Traits textbook that I am hoping to co-op for my own purposes. I am also using this as some training for the Roots class if need be.

```{r}
### I moved the data set "ProperNames Up to use on Cover"

library("traits")
library("taxize")
taxa_search(Proper_Names[1:20,]$New_Species_Name, db = "traitbank")


library("TR8")

tr8(species_list = Proper_Names[1:5,]$New_Species_Name, gui_config=TRUE, allow_persistent=TRUE)


TryNames <- read_csv(here("EnteredData", "TryAccSpecies.csv"))


Trying<-merge(Proper_Names, TryNames, by.x ="New_Species_Name", by.y = "AccSpeciesName")

unique(Trying$AccSpeciesID)

TryTraits <- read_csv(here("EnteredData", "tde20232141444.csv"))



TryTraits$count<-rowSums(TryTraits[3:length(TryTraits)] != 0)

PickedTraits<- subset(TryTraits, count>5)

colSums(PickedTraits[3:length(TryTraits)] != 0)

PickedTraits$TraitID
```



## Try Data processing
```{r}
TryData<- rtry_import(here("EnteredData", "TryData", "Traits.txt"))

rtry_trans_wider(TryData,  
                 complete.cases(TraitID))

TryData %>%
  group_by(AccSpeciesName, TraitName)
  

```

