---
title: "DCA_DataFrames"
author: "Ben J. Rivera"
date: '2023-01-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(here)
```

The goal of this script is to create and clean the necessary dataframes to be used in analysis later. Run this first then move on to other scripts.


# Reading CSVs in
Okay, this is the boring part. I need to just load in all my datasets and get everything nice and pretty. I'm going to have to deal with plant names.

## Places

Loading in data based on where it was collected

```{r}
# Ten Mile
TeMi_Coords <- read_csv(here("EnteredData", "TeMi_RelativeCoords.csv"))
TeMi_Dist <- read_csv(here("EnteredData", "TeMi_Distance.csv"))
TeMi_Cover <- read_csv(here("EnteredData", "TeMi_Cover.csv"))
TeMi_Locations <- read_csv(here("EnteredData", "TeMi_Location.csv"))


# Zmudowski
Zmud_Coords <- read_csv(here("EnteredData", "Zmud_RelativeCoords.csv"))
Zmud_Dist <- read_csv(here("EnteredData", "Zmud_Distance.csv"))
Zmud_Cover <- read_csv(here("EnteredData", "Zmud_Cover.csv"))
Zmud_Locations <- read_csv(here("EnteredData", "ZmudLocations.csv"))


# Guadalupe Nipomo
GuNi_Coords <- read_csv(here("EnteredData", "GuNi_RelativeCoords.csv"))
GuNi_Dist <- read_csv(here("EnteredData", "GuNi_Distance.csv"))
GuNi_Cover <- read_csv(here("EnteredData", "GuNi_Cover.csv"))
GuNi_Locations <- read_csv(here("EnteredData", "GuNi_Location.csv"))


```

### Create Relative Coords

```{r}
TeMi_Subset_X <- subset(TeMi_Coords, IN_FID == 1)
TeMi_Subset_Y <- subset(TeMi_Coords, IN_FID == 2)
TeMi_Sub<-merge(TeMi_Subset_X, TeMi_Subset_Y, by = "NEAR_FID")
TeMi_RelCords<- data.frame(Plot= TeMi_Sub$NEAR_FID, X = TeMi_Sub$NEAR_DIST.x, Y = TeMi_Sub$NEAR_DIST.y)
TeMi_RelCords$Plot<- merge(TeMi_RelCords[1], TeMi_Locations[1:2], by.x = "Plot", by.y = "OID_")$Comment

TeMi_RelCords<- mutate(TeMi_RelCords, X = ifelse(TeMi_RelCords$Plot == "WATERLINE",-1*X, X))

```


### Checking

```{r}
ggplot(aes(x = X, y = Y, label = Plot), data = TeMi_RelCords)+
  geom_point()+
  geom_label()

ggplot(aes(x = POINT_X, y = POINT_Y, label = Comment, color = POINT_Z), data = TeMi_Locations)+
  geom_point()+
  geom_text()+
  ggtitle("Ten Mile")
ggplot(aes(x = POINT_X, y = POINT_Y, label = Comment, color = POINT_Z), data = Zmud_Locations)+
  geom_point()+
  geom_text()+
  ggtitle("Zmudowski")
ggplot(aes(x = POINT_X, y = POINT_Y, label = Comment, color = POINT_Z), data = GuNi_Locations)+
  geom_point()+
  geom_text()+
  ggtitle("Guadalupe-Nipomo")


```

# Spanning Dataframes

## Env Data


```{r}
DuneEnvData <- read_csv(here("EnteredData", "DuneEnvData.csv"))

ggplot(aes(x= SoilTemp, y = SoilMoist, color = Site), data = DuneEnvData)+
  geom_point() #okay so site super mattered which makes sense for temp. I'll scale it

ggplot(aes(x= scale(SoilTemp), y = SoilMoist, color = Site), data = DuneEnvData)+
  geom_point() #Lol. I need to scale Soil Temp by the site! Duh! I should actually do that for Soil Moisture too but it won't be as severe


DuneEnv_scaled <- DuneEnvData %>% group_by(Site) %>% mutate(SoilTemp = scale(SoilTemp), SoilMoist=scale(SoilMoist))


DuneEnv_scaled<-plyr::ddply(DuneEnvData, "Site", transform, SoilTemp = scale(SoilTemp))
DuneEnv_scaled<-plyr::ddply(DuneEnv_scaled, "Site", transform, SoilMoist = scale(SoilMoist))


#Temp v moist
ggplot(aes(x= SoilTemp, y = SoilMoist), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) +
  geom_smooth(method = "lm", aes(color = Site), fill = NA)+
  geom_smooth(method = "lm", color = "black")

cor.test(x = DuneEnv_scaled$SoilMoist, y = DuneEnv_scaled$SoilTemp, method = "pearson") # temp and mositure are not correlated which is nice

# Slope v SoilMoist
ggplot(aes(x= Slope, y = SoilMoist), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) +
  geom_smooth(method = "lm")# No obvious pattern
cor.test(x = DuneEnv_scaled$Slope, y = DuneEnv_scaled$SoilMoist, method = "pearson") # Nutthin

#Slope v Temp ***
ggplot(aes(x= Slope , y =SoilTemp), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) +
  geom_smooth(method = "lm")
cor.test(x = DuneEnv_scaled$Slope, y = DuneEnv_scaled$SoilTemp, method = "pearson") # Topsoil, Slope explains SoilTemp

#aspect slope
ggplot(aes(x= Aspect, y = Slope), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) # No obvious pattern
summary(aov(Slope ~ Aspect, data = DuneEnv_scaled)) #nuthin


#aspect temp
ggplot(aes(x= Aspect, y = SoilTemp), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) # No obvious pattern
summary(aov(SoilTemp ~ Aspect, data = DuneEnv_scaled)) #nuthin


#aspect Moist
ggplot(aes(x= Aspect, y = SoilMoist), data = DuneEnv_scaled)+
  geom_point(aes(color = Site)) # No obvious pattern
summary(aov(SoilMoist ~ Aspect, data = DuneEnv_scaled)) #nuthin
```




## Phenology

```{r}
Pheno <-read_csv(here("EnteredData", "DunePhenoTrait.csv"))

head(Pheno)
```

