---
title: "1 Within Plot Level"
author: "Ben J. Rivera"
date: '2023-01-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(vegan)
library(interactions)
library(corrplot)
library(marginaleffects)
library(lme4)
library(ggstance)
library(ggridges)
library(lattice)

load(file = "DuneComm1.RData") #this allows me to use the global data. So I can make dataframes elsewhere and then use them here.

```

Okay, the goal of this section is all about understanding how a much plant variation there can be within a 2x2 m quadrat. The goal is to compare the focal communities to their immediate surroundings. If they are tightly conserved, that would suggest a niche-y assembly which will then lead to an examination of which environmental variables matter. If no environmental variables are good, then we either did not measure the correct ones or we can blame spatial auto-correlation (See: 3_WithinSite/BetweenPointsLevel). If it is spatial auto-correlation and/or random, then we are thinking about neutral. I expect the relative importance of the environemnt to vary with environmental stress.

# *Species Composition*
## Ten Mile

This is a section for analyzing data only from ten mile. I will repeat same process for each location.

```{r}
#I make unique identifiers for each plot
TeMi_Names<- colnames(TeMi_Cover)
TeMi_Names<-gsub('.{1}$', '', TeMi_Names)
TeMi_Names<-unique(gsub('TeMi.', '', TeMi_Names))
TeMi_Diss<-data.frame(Plot = rep("", length(TeMi_Names)), Dissimilarity= rep(0, length(TeMi_Names)))


for (i in 1:length(TeMi_Names)) {
  Temp_Cover<-TeMi_Cover %>% select(starts_with(paste0("TeMi.", TeMi_Names[i])))
  #Temp_Cover<- cbind(Temp_Cover[1], rowSums(Temp_Cover[2:5], na.rm = TRUE)/4)
  Temp_Diss<-vegdist(t(Temp_Cover), na.rm = TRUE)
  Temp_Av_Dis<-sum(as.matrix(Temp_Diss)[,1])/4
  TeMi_Diss$Plot[i]<- TeMi_Names[i]
  TeMi_Diss$Dissimilarity[i]<- ifelse(is.na(Temp_Av_Dis) == TRUE, 0, ifelse(Temp_Av_Dis == "NaN", 0, Temp_Av_Dis))

}



densityplot(TeMi_Diss$Dissimilarity)

TeMi_Diss$Plot<- as.numeric(TeMi_Diss$Plot)

TeMi_Diss<-merge(TeMi_Diss, TeMi_RelCords)

TeMi_Diss<-merge(TeMi_Diss, subset(DuneEnv_scaled, Site == "TeMi"))


TeMi_pcor<- TeMi_Diss %>%
  select(-c(Site, Plot, Aspect, Date, Collectors, dry_weight, Notes))%>%
  as.matrix()%>%
  cor.mtest(use =  "complete")

TeMi_Diss %>%
  select(-c(Site, Plot, Aspect, Date, Collectors, dry_weight, Notes))%>%
  as.matrix()%>%
  cor(use =  "complete")%>%
  corrplot( type = "upper", order = "hclust",  tl.col = "black", tl.srt = 45, p.mat = TeMi_pcor$p)
```

### Graphs and models

#### Distance

```{r}
densityplot(TeMi_Diss$X)
densityplot(TeMi_Diss$Y)
densityplot(TeMi_Diss$Z)

summary(lm(Dissimilarity ~ X, data = TeMi_Diss)) # Distance from waterline not sig
ggplot(aes(x = X, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()+
  labs(x = "X-Coordinate", y = "Compositional Dissimilarity", title = "Ten-Mile (North)")

summary(lm(Dissimilarity ~ Y, data = TeMi_Diss))
ggplot(aes(x = Y, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ Z, data = TeMi_Diss))
ggplot(aes(x = Z, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()


summary(lm(Dissimilarity ~ X+Y+Z, data = TeMi_Diss))


```

#### Environmental

```{r}
densityplot(TeMi_Diss$SoilTemp)
densityplot(TeMi_Diss$SoilMoist)
densityplot(TeMi_Diss$Slope)
densityplot(TeMi_Diss$pH)
densityplot(TeMi_Diss$cond_us.cm)
densityplot(TeMi_Diss$salinity_ppt)



summary(lm(Dissimilarity ~ SoilTemp+SoilMoist+Slope+Aspect+pH, data = TeMi_Diss))

summary(lm(Dissimilarity ~ SoilTemp, data = TeMi_Diss))
ggplot(aes(x = SoilTemp, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()


summary(lm(Dissimilarity ~ SoilMoist, data = TeMi_Diss))
ggplot(aes(x = SoilMoist, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ SoilTemp*SoilMoist, data = TeMi_Diss))

summary(lm(Dissimilarity ~ Slope, data = TeMi_Diss))
ggplot(aes(x = Slope, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()


summary(lm(Dissimilarity ~ Aspect, data = TeMi_Diss))
ggplot(aes(x = Aspect, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()

######## INTERESTING! ######
summary(lm(Dissimilarity ~ pH, data = TeMi_Diss))
ggplot(aes(x = pH, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ cond_us.cm, data = TeMi_Diss))
ggplot(aes(x = cond_us.cm, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()+
  geom_smooth(method = "lm")

summary(lm(Dissimilarity ~ salinity_ppt, data = TeMi_Diss))
ggplot(aes(x = salinity_ppt, y = Dissimilarity), data = TeMi_Diss)+
  geom_point()+
  geom_smooth(method = "lm")

```
#### Combined

```{r}
cor.test(x = TeMi_Diss$X, y = TeMi_Diss$salinity_ppt, method = "pearson")
ggplot(aes(x = X, y = salinity_ppt), data = TeMi_Diss)+ #Weird that X isn't correlated with dissim
  geom_point()+
  geom_smooth(method = "lm")

mod.x.temi<-lm(Dissimilarity ~ salinity_ppt*X, data = TeMi_Diss)
summary(mod.x.temi) #WHAT!?!?!
interact_plot(mod.x.temi, pred = X, modx = salinity_ppt)

```

#### Negligible effects

```{r}

TeMi.mod.1<- lm(Dissimilarity ~ SoilTemp + SoilMoist + Slope + Aspect + X + Z + pH, data = TeMi_Diss)
summary(TeMi.mod.1)
TeMi.Null.1 <- lm(Dissimilarity ~ 1, data = TeMi_Diss)
anova(TeMi.mod.1, TeMi.Null.1)

TeMi.pre.1<- predictions(TeMi.mod.1)

TeMi.cmp<- avg_comparisons(TeMi.mod.1,  variables =  list(SoilTemp = "minmax", SoilMoist = "minmax", Slope = "minmax", Aspect = "pairwise", X = "minmax", Z = "minmax", pH = "minmax"), conf_level = 0.90, type = "response")


TeMi.cmp<-TeMi.cmp %>%
  group_by(term) %>%
  summarise_all(.funs = mean)

ggplot(TeMi.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1.5)+
  geom_errorbarh() + 
  geom_point(size = 2) +
  theme_classic()


```




## Zmudowski
```{r}
#I make unique identifiers for each plot
Zmud_Names<- colnames(Zmud_Cover)
Zmud_Names<-gsub('.{1}$', '', Zmud_Names)
Zmud_Names<-unique(gsub('Zmud.', '', Zmud_Names))
Zmud_Diss<-data.frame(Plot = rep("", length(Zmud_Names)), Dissimilarity= rep(0, length(Zmud_Names)))


for (i in 1:length(Zmud_Names)) {
  Temp_Cover<-Zmud_Cover %>% select(starts_with(paste0("Zmud.", Zmud_Names[i])))
  #Temp_Cover<- cbind(Temp_Cover[1], rowSums(Temp_Cover[2:5], na.rm = TRUE)/4)
  Temp_Diss<-vegdist(t(Temp_Cover), na.rm = TRUE)
  Temp_Av_Dis<-sum(as.matrix(Temp_Diss)[,1])/4
  Zmud_Diss$Plot[i]<- Zmud_Names[i]
  Zmud_Diss$Dissimilarity[i]<- ifelse(is.na(Temp_Av_Dis) == TRUE, 0, ifelse(Temp_Av_Dis == "NaN", 0, Temp_Av_Dis))

}


densityplot(Zmud_Diss$Dissimilarity)

Zmud_Diss$Plot<- as.numeric(Zmud_Diss$Plot)

Zmud_Diss<-merge(Zmud_Diss, Zmud_RelCords)


Zmud_Diss<-merge(Zmud_Diss, subset(DuneEnv_scaled, Site == "Zmud"))

Zmud_pcor<- Zmud_Diss %>%
  select(-c(Site, Plot, Aspect, Date, Collectors, dry_weight, Notes, pH, cond_us.cm, salinity_ppt))%>%
  as.matrix()%>%
  cor.mtest(use =  "complete")

Zmud_Diss %>%
  select(-c(Site, Plot, Aspect, Date, Collectors, dry_weight, Notes, pH, cond_us.cm, salinity_ppt))%>%
  as.matrix()%>%
  cor(use =  "complete") %>%
  corrplot(type = "upper", order = "hclust", tl.col = "black", tl.srt = 45, p.mat = Zmud_pcor$p)
```

### Graphs and models

#### Distance
```{r}


densityplot(Zmud_Diss$X)
densityplot(Zmud_Diss$Y)
densityplot(Zmud_Diss$Z)

summary(lm(Dissimilarity ~ X, data = Zmud_Diss)) # Distance from waterline SIG HERE!
ggplot(aes(x = X, y = Dissimilarity), data = Zmud_Diss)+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(x = "X-Coordinate", y = "Compositional Dissimilarity", title = "Zmudowski (middle)")

 summary(lm(Dissimilarity ~ Y, data = Zmud_Diss))
ggplot(aes(x = Y, y = Dissimilarity), data = Zmud_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ Z, data = Zmud_Diss))
ggplot(aes(x = Z, y = Dissimilarity), data = Zmud_Diss)+
  geom_point()


summary(lm(Dissimilarity ~ X*Y*Z, data = Zmud_Diss))


```
#### Environmental

```{r}
densityplot(Zmud_Diss$SoilTemp)
densityplot(Zmud_Diss$SoilMoist)
densityplot(Zmud_Diss$Slope)

summary(lm(Dissimilarity ~ SoilTemp+SoilMoist+Slope+Aspect, data = Zmud_Diss))

summary(lm(Dissimilarity ~ SoilTemp, data = Zmud_Diss))
ggplot(aes(x = SoilTemp, y = Dissimilarity), data = Zmud_Diss)+
  geom_point()+
  geom_smooth(method = "lm")


summary(lm(Dissimilarity ~ log(2+SoilMoist), data = Zmud_Diss))
ggplot(aes(x = log(2+SoilMoist), y = Dissimilarity), data = Zmud_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ SoilTemp*SoilMoist, data = Zmud_Diss))

summary(lm(Dissimilarity ~ Slope, data = Zmud_Diss))
ggplot(aes(x = Slope, y = Dissimilarity), data = Zmud_Diss)+
  geom_point()


summary(lm(Dissimilarity ~ Aspect, data = Zmud_Diss))
ggplot(aes(x = Aspect, y = Dissimilarity), data = Zmud_Diss)+
  geom_point()
```
#### Combined

```{r}
summary(lm(Dissimilarity ~ SoilTemp*X, data = Zmud_Diss))
summary(lm(Dissimilarity ~ SoilMoist*X, data = Zmud_Diss))

summary(lm(Dissimilarity ~ SoilTemp*Y, data = Zmud_Diss))
summary(lm(Dissimilarity ~ SoilMoist*Y, data = Zmud_Diss))

summary(lm(Dissimilarity ~ SoilTemp*Z, data = Zmud_Diss))
summary(lm(Dissimilarity ~ SoilMoist*Z, data = Zmud_Diss))

```


#### Negligible effects

```{r}

Zmud.mod.1<- lm(Dissimilarity ~ SoilTemp + SoilMoist + Slope + Aspect + X + Z, data = Zmud_Diss)
summary(Zmud.mod.1)
Zmud.Null.1 <- lm(Dissimilarity ~ 1, data = Zmud_Diss)
anova(Zmud.mod.1, Zmud.Null.1)


Zmud.cmp<- avg_comparisons(Zmud.mod.1,  variables =  list(SoilTemp = "minmax", SoilMoist = "minmax", Slope = "minmax", Aspect = "pairwise", X = "minmax", Z = "minmax"), conf_level = 0.90, type = "response")


Zmud.cmp<-Zmud.cmp %>%
  group_by(term) %>%
  summarise_all(.funs = mean)

ggplot(Zmud.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1.5)+
  geom_errorbarh() + 
  geom_point(size = 2) +
  theme_classic()


```
## Guadalupe-Nipomo
```{r}
#I make unique identifiers for each plot
GuNi_Names<- colnames(GuNi_Cover)
GuNi_Names<-gsub('.{1}$', '', GuNi_Names)
GuNi_Names<-unique(gsub('GuNi.', '', GuNi_Names))
GuNi_Diss<-data.frame(Plot = rep("", length(GuNi_Names)), Dissimilarity= rep(0, length(GuNi_Names)))


for (i in 1:length(GuNi_Names)) {
  Temp_Cover<-GuNi_Cover %>% select(starts_with(paste0("GuNi.", GuNi_Names[i])))
  #Temp_Cover<- cbind(Temp_Cover[1], rowSums(Temp_Cover[2:5], na.rm = TRUE)/4)
  Temp_Diss<-vegdist(t(Temp_Cover), na.rm = TRUE)
  Temp_Av_Dis<-sum(as.matrix(Temp_Diss)[,1])/4
  GuNi_Diss$Plot[i]<- GuNi_Names[i]
  GuNi_Diss$Dissimilarity[i]<- ifelse(is.na(Temp_Av_Dis) == TRUE, 0, ifelse(Temp_Av_Dis == "NaN", 0, Temp_Av_Dis))

}


densityplot(GuNi_Diss$Dissimilarity)

GuNi_Diss$Plot<- as.numeric(GuNi_Diss$Plot)

GuNi_Diss<-merge(GuNi_Diss, GuNi_RelCords)

GuNi_Diss<-merge(GuNi_Diss, subset(DuneEnv_scaled, Site == "GuNi"))


GuNi_pcor<- GuNi_Diss %>%
  select(-c(Site, Plot, Aspect, Date, Collectors, dry_weight, Notes))%>%
  as.matrix()%>%
  cor.mtest(use =  "complete")


GuNi_Diss %>%
  select(-c(Site, Plot, Aspect, Date, Collectors, dry_weight, Notes))%>%
  as.matrix()%>%
  cor(use =  "complete")%>%
  corrplot( type = "upper", order = "hclust",  tl.col = "black", tl.srt = 45, p.mat = GuNi_pcor$p)


```

### Graphs and models

#### Distance
```{r}
densityplot(GuNi_Diss$X)
densityplot(GuNi_Diss$Y)
densityplot(GuNi_Diss$Z)

summary(lm(Dissimilarity ~ X, data = GuNi_Diss)) # Distance from waterline not sig
ggplot(aes(x = X, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()+
  labs(x = "X-Coordinate", y = "Compositional Dissimilarity", title = "Guadalupe-Nipomo (South)")

summary(lm(Dissimilarity ~ Y, data = GuNi_Diss))
ggplot(aes(x = Y, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ Z, data = GuNi_Diss))
ggplot(aes(x = Z, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()


summary(lm(Dissimilarity ~ X+Y+Z,  data = GuNi_Diss))


```



#### Environmental

```{r}
densityplot(GuNi_Diss$SoilTemp)
densityplot(GuNi_Diss$SoilMoist)
densityplot(GuNi_Diss$Slope)
densityplot(GuNi_Diss$pH)
densityplot(GuNi_Diss$cond_us.cm)
densityplot(GuNi_Diss$salinity_ppt)

```


```{r}

summary(lm(Dissimilarity ~ SoilTemp+SoilMoist+Slope+Aspect+pH+salinity_ppt, data = GuNi_Diss))

summary(lm(Dissimilarity ~ SoilTemp, data = GuNi_Diss))
ggplot(aes(x = SoilTemp, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()


summary(lm(Dissimilarity ~ SoilMoist, data = GuNi_Diss))
ggplot(aes(x = SoilMoist, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ SoilTemp*SoilMoist, data = GuNi_Diss))

summary(lm(Dissimilarity ~ Slope, data = GuNi_Diss))
ggplot(aes(x = Slope, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()


summary(lm(Dissimilarity ~ Aspect, data = GuNi_Diss))
ggplot(aes(x = Aspect, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ pH, data = GuNi_Diss))
ggplot(aes(x = pH, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ cond_us.cm, data = GuNi_Diss))
ggplot(aes(x = cond_us.cm, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()

summary(lm(Dissimilarity ~ salinity_ppt, data = GuNi_Diss))
ggplot(aes(x = salinity_ppt, y = Dissimilarity), data = GuNi_Diss)+
  geom_point()


```


#### Negligible effects

```{r}

GuNi.mod.1<- lm(Dissimilarity ~ SoilTemp + SoilMoist + Slope + Aspect + X + Z + pH, data = subset(GuNi_Diss, is.na(pH) == FALSE))
summary(GuNi.mod.1)
GuNi.Null.1 <- lm(Dissimilarity ~ 1, data = subset(GuNi_Diss, is.na(pH) == FALSE))
anova(GuNi.mod.1, GuNi.Null.1)

GuNi.pre.1<- predictions(GuNi.mod.1)

GuNi.cmp<- avg_comparisons(GuNi.mod.1,  variables =  list(SoilTemp = "minmax", SoilMoist = "minmax", Slope = "minmax", Aspect = "pairwise", X = "minmax", Z = "minmax", pH = "minmax"), conf_level = 0.90, type = "response")


GuNi.cmp<-GuNi.cmp %>%
  group_by(term) %>%
  summarise_all(.funs = mean)

ggplot(GuNi.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1.5)+
  geom_errorbarh() + 
  geom_point(size = 2) +
  theme_classic()


```

# Mixed

```{r}
Comb_Dis<-rbind(TeMi_Diss, Zmud_Diss, GuNi_Diss)

Comb_pcor<- Comb_Dis %>%
  select(-c(Site, Plot, Aspect, Date, Collectors, dry_weight, Notes))%>%
  as.matrix()%>%
  cor.mtest(use =  "complete")


Comb_Dis %>%
  select(-c(Site, Plot, Aspect, Date, Collectors, dry_weight, Notes))%>%
  as.matrix()%>%
  cor(use =  "complete")%>%
  corrplot( type = "upper", order = "hclust",  tl.col = "black", tl.srt = 45, p.mat = GuNi_pcor$p, diag = FALSE)
```

```{r}

densityplot(Comb_Dis$Dissimilarity)

densityplot(subset(Comb_Dis, Dissimilarity >0)$Dissimilarity)




summary(lm(Dissimilarity ~ X, data = Comb_Dis)) # Distance from waterline not sig
ggplot(aes(x = X, y = Dissimilarity), data = Comb_Dis)+
  geom_point()+
  labs(x = "X-Coordinate", y = "Compositional Dissimilarity", title = "Guadalupe-Nipomo (South)")

summary(lm(Dissimilarity ~ Y, data = Comb_Dis))
ggplot(aes(x = Y, y = Dissimilarity), data = Comb_Dis)+
  geom_point()

summary(lm(Dissimilarity ~ Z, data = Comb_Dis))
ggplot(aes(x = Z, y = Dissimilarity), data = Comb_Dis)+
  geom_point()


summary(lm(Dissimilarity ~ X+Y+Z,  data = Comb_Dis))

```


```{r}

summary(lm(Dissimilarity ~ SoilTemp+SoilMoist+Slope+Aspect+pH+salinity_ppt, data = Comb_Dis))

summary(lm(Dissimilarity ~ SoilTemp, data = Comb_Dis))
ggplot(aes(x = SoilTemp, y = Dissimilarity), data = Comb_Dis)+
  geom_point()


summary(lm(Dissimilarity ~ SoilMoist, data = Comb_Dis))
ggplot(aes(x = SoilMoist, y = Dissimilarity), data = Comb_Dis)+
  geom_point()

summary(lm(Dissimilarity ~ SoilTemp*SoilMoist, data = Comb_Dis))

summary(lm(Dissimilarity ~ Slope, data = Comb_Dis))
ggplot(aes(x = Slope, y = Dissimilarity), data = Comb_Dis)+
  geom_point()


summary(lm(Dissimilarity ~ Aspect, data = Comb_Dis))
ggplot(aes(x = Aspect, y = Dissimilarity), data = Comb_Dis)+
  geom_point()

summary(lm(Dissimilarity ~ pH, data = Comb_Dis))
ggplot(aes(x = pH, y = Dissimilarity), data = Comb_Dis)+
  geom_point()

summary(lm(Dissimilarity ~ cond_us.cm, data = Comb_Dis))
ggplot(aes(x = cond_us.cm, y = Dissimilarity), data = Comb_Dis)+
  geom_point()

summary(lm(Dissimilarity ~ salinity_ppt, data = Comb_Dis)) #***
ggplot(aes(x = salinity_ppt, y = Dissimilarity), data = Comb_Dis)+
  geom_point()+
  geom_smooth(method = "lm")


```

### Negligible Effects

```{r}
Mixed.mod.1 <- lmer(Dissimilarity ~ SoilTemp + SoilMoist + Slope + Aspect +  X +  Z  + (1|Site), data = subset(Comb_Dis, Dissimilarity >0.0))
summary(Mixed.mod.1)
Mixed.Null.1 <- lmer(Dissimilarity ~ 1  + (1|Site), data = subset(Comb_Dis, Dissimilarity >0.0))
Mixed.ST<- lmer(Dissimilarity ~ SoilTemp  + (1|Site), data = subset(Comb_Dis, Dissimilarity >0.0))
Mixed.X<- lmer(Dissimilarity ~ X  + (1|Site), data = subset(Comb_Dis, Dissimilarity >0.0))

anova(Mixed.mod.1, Mixed.Null.1)
anova(Mixed.ST, Mixed.Null.1)
anova(Mixed.X, Mixed.Null.1)
anova(Mixed.ST, Mixed.X)

Mixed.cmp<- avg_comparisons(Mixed.mod.1, variables = list(SoilTemp = "minmax", SoilMoist = "minmax", Slope = "minmax", Aspect = "pairwise", X = "minmax", Z = "minmax"),   conf_level = 0.90, type = "response")

Mixed.cmp <- as.data.frame(Mixed.cmp)


for(j in 1:dim(Mixed.cmp)[1]){
  if(Mixed.cmp$term[j] == "Aspect"){
    if(Mixed.cmp$estimate[j] <= 0){
   Mixed.cmp$estimate[j] <- (Mixed.cmp$estimate[j] * -1)
   NewLow <- (Mixed.cmp$conf.high[j] * -1)
   Mixed.cmp$conf.high[j] <- (Mixed.cmp$conf.low[j] * -1)
   Mixed.cmp$conf.low[j] <- NewLow
  } 
}
}



ggplot(Mixed.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term,
                 group = contrast)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1.5)+
  geom_errorbarh(aes(xmin = conf.low,  xmax = conf.high,),position = position_dodgev(height = 1.5)) + 
  geom_point(size = 2, position = position_dodgev(height = 1.5)) +
  theme_classic()


Mixed.cmp<-Mixed.cmp %>%
  group_by(term) %>%
  summarise_all(.funs = mean)



ggplot(Mixed.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term,
                 group = contrast)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1.5)+
  geom_errorbarh(aes(xmin = conf.low,  xmax = conf.high,),position = position_dodgev(height = 1.5)) + 
  geom_point(size = 4, position = position_dodgev(height = 1.5)) +
  theme_classic()+
  labs(x = "Estimated Change in Community Dissimilarity", y = "Term")+
  theme(text = element_text(family = "serif", size = 18))

ggsave("Within-Marginal.png", dpi = 600, height = 7, width = 10)

plot_predictions(Mixed.mod.1, condition = "SoilTemp")+
  geom_point(data = subset(Comb_Dis, Dissimilarity >0), size = 3,aes(x = SoilTemp, y = Dissimilarity, color = Site, shape = Site))+
  labs(x = "Soil Temperature (°C)", y =  "Community Dissimilarity")+
  scale_color_manual(labels = c("Guadalupe-Nipomo", "Ten Mile", "Zmudowski"), values = c("#d55e00", "#cc79a7", "#009e73"), name = "Site")+
  scale_shape_manual(labels = c("Guadalupe-Nipomo", "Ten Mile", "Zmudowski"), values = c(1,2,0), name = "Site")+
  theme_classic()+
  theme(text = element_text(family = "serif", size = 18))
  #geom_smooth(data = subset(Comb_Dis, Dissimilarity >0), aes(x = SoilTemp, y = Dissimilarity), method = "lm")

ggsave("Within-SoilTemp.png", dpi = 600, height = 7, width = 10)


```

#### Randomization Test

```{r}
Tracker<- data.frame()
SingleTrack<-0

for (i in 1:100) {
  
  Rand.1<- transform(Comb_Dis, Dissimilarity = sample(Dissimilarity) )
  Rand.mod.1 <- lmer(Dissimilarity ~ SoilTemp + SoilMoist + Slope + Aspect + X +  Z  + (1|Site), data = subset(Rand.1, Dissimilarity >0))
  
  
  while(isSingular(Rand.mod.1) == TRUE){
    SingleTrack <- SingleTrack + 1
   Rand.1<- transform(Comb_Dis, Dissimilarity = sample(Dissimilarity) )
  Rand.mod.1 <- lmer(Dissimilarity ~ SoilTemp + SoilMoist + Slope + Aspect + X +  Z  + (1|Site), data = subset(Rand.1, Dissimilarity >0))
    } 
  
    Rand.cmp<- avg_comparisons(Rand.mod.1, variables = list(SoilTemp = "minmax", SoilMoist = "minmax", Slope = "minmax", Aspect = "pairwise", X = "minmax", Z = "minmax"),   conf_level = 0.90, type = "response")
      
  for(j in 1:dim(Rand.cmp)[1]){
  if(Rand.cmp$term[j] == "Aspect"){
    if(Rand.cmp$estimate[j] <= 0){
   Rand.cmp$estimate[j] <- (Rand.cmp$estimate[j] * -1)
   NewLow <- (Rand.cmp$conf.high[j] * -1)
   Rand.cmp$conf.high[j] <- (Rand.cmp$conf.low[j] * -1)
   Rand.cmp$conf.low[j] <- NewLow
  } 
}
}
  
  Rand.cmp<-Rand.cmp %>%
    group_by(term) %>%
    summarise_all(.funs = mean)

   Run <- rep(i, dim(Rand.cmp)[1])
    Rand.cmp<- cbind(Rand.cmp, Run)


    
    
    Tracker <- rbind(Tracker, Rand.cmp)
    print(i)
    
  
}


    # ggplot(Tracker, aes(x = estimate,
    #                xmin = conf.low,
    #                xmax = conf.high,
    #                y = term,
    #                group = contrast)) +
    #   annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
    #   geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") +
    #   geom_vline(xintercept = 0, linetype = "solid", linewidth = 1.5)+
    #   geom_errorbarh(position = position_dodgev(height = 1.5)) +
    #   geom_point(size = 2, position = position_dodgev(height = 1.5)) +
    #   theme_classic()
    
ggplot(Tracker, aes(x = estimate, 
                    y = term, 
                    group = term))+
  geom_density_ridges(fill = "blue", alpha = 0.75)+
      annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
      geom_vline(xintercept = c(-0.1, .1), linetype = "dashed")

length(unique(Tracker$Run))

Tracker %>% group_by(term) %>% summarize_all(mean)
```


```{r}
save.image(file="DuneComm2.RData") 
```


