---
title: "2_BetweenPlotsLevel"
author: "Ben J. Rivera"
date: "2023-07-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(vegan)
library(interactions)
#devtools::install_github("vmikk/metagMisc")
library("metagMisc") # for dist2list command
library(circular) # to calculate dist for aspect
library(marginaleffects)
library(ggstance)
library(lattice)
library(ggpubr)
library(cowplot)
library(extrafont)
library(lme4)
library(modelsummary)


load(file = "DuneComm3.RData") #this allows me to use the global data. So I can make dataframes elsewhere and then use them here.
CircleConverter <- data.frame(Aspect = c("N", "NW","W","SW", "S", "SE","E","NE"), 
                     Angle = c(0,45,90,135,180,225,270,315)) #critical for conversion of aspect


```

The goal of this file is look at the the between plot level of our study. As we did in previous files, we begin with measuring each site individually. For the analysis more relevant to the manuscript, scroll to the heading title "**Combined Analysis**" where we grouped each site together. From there, we created our models and figures. 

# Ten Mile

## Cover

```{r}
TeMi2_Cover <- data.frame()

for (i in 1:length(TeMi_Names)) { #for-loop to combine focal and concentric quadrats 
  Temp_Cover<-TeMi_Cover %>% select(starts_with(paste0("TeMi.", TeMi_Names[i])))
  Temp_Concentric<-rowSums(Temp_Cover[2:5]/4) #CHECK THIS
  Temp_Cover2<- (Temp_Cover[1]*.25)+(Temp_Concentric*.75) #CHECK THIS
  if(i ==1) {
    TeMi2_Cover<- Temp_Cover2
  } else {
    TeMi2_Cover<-cbind(TeMi2_Cover, Temp_Cover2)
  }
}

mean(colSums(TeMi_Cover, na.rm = TRUE))
mean(colSums(TeMi2_Cover, na.rm = TRUE)) #I think these numbers lining up means it worked right
```


```{r}

#TeMi2_Cover <- TeMi2_Cover[,colSums(is.na(TeMi2_Cover))<nrow(TeMi2_Cover)]
TeMi2_Cover[is.na(TeMi2_Cover)] <- 0

colSums(TeMi2_Cover) > 100

bareground<- vector(mode="double", length=ncol(TeMi2_Cover))
names(bareground) <- "bareground"

for(l in 1:length(TeMi2_Cover)) {
  bareground[l] <- 100-sum(TeMi2_Cover[l])
}
TeMi2_Cover <- rbind(TeMi2_Cover, bareground)

rownames(TeMi2_Cover)[dim(TeMi2_Cover)[1]] <- "bareground"
```


```{r}

TeMi2_Dis<-vegdist(t(TeMi2_Cover), na.rm = TRUE, method = "bray") #Calculate Dissim between plots

TeMi2_CoverNMDS<-metaMDS(comm =t(TeMi2_Cover), distance = "bray"  ) #just doing a gross NMDS for visualization
ordiplot(TeMi2_CoverNMDS, type = "text")

TeMi2<- dist2list(TeMi2_Dis, tri = TRUE)
densityplot(TeMi2$value) 
densityplot(sqrt(TeMi2$value)) 

```
I'm not sure if adding in bareground was smart or not

## Environmental

```{r}
TeMi2_Env<-TeMi_Diss[c(1,7:9 , 14:16)]
TeMi2_Env<- data.frame(TeMi2_Env[,-1], row.names=TeMi2_Env[,1])

TeMi2_Env_dissim <- vegdist(TeMi2_Env, method = "gower")
TeMi2_Env_dissim_long<- dist2list(TeMi2_Env_dissim, tri = TRUE)
TeMi2_EnvNMDS<-metaMDS(comm = TeMi2_Env, distance = "gower")
plot(TeMi2_EnvNMDS)
densityplot(TeMi2_Env_dissim_long$value)
densityplot(TeMi2_Env_dissim_long$value*.8)


```

## Aspect

Aspect has been such a fun puzzle to solve. Ultimately, I have decided on the two dissimilarity matrices route.

```{r}

TeMi_Aspect<- data.frame(TeMi_Diss$Aspect, row.names = TeMi_Diss[,1])


for(a in 1:nrow(CircleConverter)){
  TeMi_Aspect$TeMi_Diss.Aspect[which(TeMi_Aspect == CircleConverter$Aspect[a])] <- CircleConverter$Angle[a]

  
}

TeMi_Aspect$TeMi_Diss.Aspect <- as.numeric(TeMi_Aspect$TeMi_Diss.Aspect)
TeMi_Circle<- circular(TeMi_Aspect, units = "degrees", modulo = "2pi", rotation = "clock", template = "geographics")
plot.circular(TeMi_Circle, stack = TRUE)
TeMi_CircleDiss<-dist.circular(TeMi_Circle, method = "geodesic")/pi #well this actually make the most sense
TeMi_CircleDissLong<- dist2list(TeMi_CircleDiss, tri = TRUE)
densityplot(TeMi_CircleDissLong$value)
summary(TeMi_CircleDissLong$value)
```

## Env Combined

```{r}
which(TeMi2_Env_dissim_long$col != TeMi_CircleDissLong$col)
which(TeMi2_Env_dissim_long$row != TeMi_CircleDissLong$row)

TeMi2$EnvDis<-(TeMi2_Env_dissim_long$value * 0.8) + (TeMi_CircleDissLong$value * 0.2)
which(TeMi2_Env_dissim_long$value == TeMi2$EnvDis)

densityplot(TeMi2$EnvDis)
summary(TeMi2$EnvDis)
summary(TeMi2$value)


ggplot(aes(x = EnvDis, y = value), data = TeMi2)+
  geom_point()+
  geom_smooth(method = "lm")

ggplot(aes(x = EnvDis, y = sqrt(value)), data = TeMi2)+
  geom_point()+
  geom_smooth(method = "lm")

summary(lm(value ~ EnvDis, data = TeMi2))
summary(lm(sqrt(value) ~ EnvDis, data = TeMi2))
```

## Actual Distances

```{r}


TeMi_difvalues<-setdiff(TeMi_RelCords$Plot, as.character(TeMi_Diss$Plot))



`%not_in%` <- purrr::negate(`%in%`)
TeMi2_RelCord<-TeMi_RelCords[TeMi_RelCords$Plot %not_in% TeMi_difvalues,]


TeMi2_RelCord <- data.frame(TeMi2_RelCord[,-1], row.names=TeMi2_RelCord[,1])
TeMi2_RelCord
TeMi2_X<-designdist(TeMi2_RelCord[1], method = "A-B", terms = "minimum")

TeMi2_Y<-designdist(TeMi2_RelCord[2], method = "A-B", terms = "minimum")
TeMi2_Z<-designdist(TeMi2_RelCord[3], method = "A-B", terms = "minimum")


TeMi2_Dist<-data.frame(dist2list(TeMi2_X, tri = TRUE)[1:2],Xdist = dist2list(TeMi2_X, tri = TRUE)$value, Ydist = dist2list(TeMi2_Y, tri = TRUE)$value, Zdist = dist2list(TeMi2_Z, tri = TRUE)$value)

TeMi2_Dist$TotalDist<-sqrt(TeMi2_Dist$Xdist^2 + TeMi2_Dist$Ydis^2 + TeMi2_Dist$Zdist^2)


densityplot(TeMi2_Dist$Xdist)
densityplot(TeMi2_Dist$Ydist)
densityplot(TeMi2_Dist$Zdist)
densityplot(TeMi2_Dist$TotalDist)

TeMi2<-data.frame(TeMi2_Dist[1:2], Dissim = TeMi2$value, TeMi2[4], TeMi2_Dist[3:6])


```

## Plots and models

```{r}

TeMi_EnvDis.mod <-lm(Dissim ~ EnvDis, data = TeMi2)
summary(TeMi_EnvDis.mod) # .

ggplot(aes(y = Dissim, x = EnvDis ), data = TeMi2)+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Ten-Mile: Environmental Similarity between Points")


summary(lm(Dissim ~ abs(Xdist), data = TeMi2))

ggplot(aes(y = Dissim, x = abs(Xdist) ), data = TeMi2)+
  geom_point()


summary(lm(Dissim ~ abs(Ydist), data = TeMi2))

ggplot(aes(y = Dissim, x = abs(Ydist) ), data = TeMi2)+
  geom_point()

summary(lm(Dissim ~ sqrt(abs(Zdist)), data = TeMi2)) # **

ggplot(aes(y = Dissim, x = sqrt(abs(Zdist)) ), data = TeMi2)+
  geom_point()+
  geom_smooth(method = "lm")


TeMi_TotalDist.mod<- lm(Dissim ~ TotalDist, data = TeMi2)
summary(TeMi_TotalDist.mod) # .

ggplot(aes(y = Dissim, x = TotalDist ), data = TeMi2)+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Ten-Mile: Distance between Points")


anova(TeMi_EnvDis.mod, TeMi_TotalDist.mod, test = "Chisq")

summary(lm(Dissim ~ EnvDis+TotalDist, data = TeMi2)) # **
summary(lm(Dissim ~ EnvDis*TotalDist, data = TeMi2)) # .

```



## Marginal Effects

```{r}
TeMi_Combined.mod <- lm(Dissim ~ TotalDist+EnvDis, data = TeMi2)
summary(TeMi_Combined.mod)
TeMi.pre<-predictions(TeMi_Combined.mod)
ggplot(aes(x = EnvDis, y = estimate), data = TeMi.pre)+
  geom_point()


TeMi.cmp<- avg_comparisons(TeMi_Combined.mod,  variables =  list(TotalDist = "minmax", EnvDis = "minmax"), conf_level = 0.90, type = "response")

TeMi.cmp

ggplot(TeMi.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", size = 1.5)+
  geom_errorbarh() + 
  geom_point(size = 2) +
  theme_classic()+
  labs(title = "Ten Mile")

plot_comparisons(TeMi_Combined.mod,
                 variables = "EnvDis",
                 condition = "TotalDist",
                 comparison = "ratio") 
```





# Zmudowski



## Cover

```{r}
Zmud2_Cover <- data.frame()

for (i in 1:length(Zmud_Names)) { #for-loop to combine focal and concentric quadrats 
  Temp_Cover<-Zmud_Cover %>% select(starts_with(paste0("Zmud.", Zmud_Names[i])))
  Temp_Concentric<-rowSums(Temp_Cover[2:5]/4) #CHECK THIS
  Temp_Cover2<- (Temp_Cover[1]*.25)+(Temp_Concentric*.75) #CHECK THIS
  if(i ==1) {
    Zmud2_Cover<- Temp_Cover2
  } else {
    Zmud2_Cover<-cbind(Zmud2_Cover, Temp_Cover2)
  }
}

mean(colSums(Zmud_Cover, na.rm = TRUE))
mean(colSums(Zmud2_Cover, na.rm = TRUE)) #I think these numbers lining up means it worked right
```


```{r}

#Zmud2_Cover <- Zmud2_Cover[,colSums(is.na(Zmud2_Cover))<nrow(Zmud2_Cover)]
Zmud2_Cover[is.na(Zmud2_Cover)] <- 0

colSums(Zmud2_Cover) > 100

bareground<- vector(mode="double", length=ncol(Zmud2_Cover))
names(bareground) <- "bareground"

for(l in 1:length(Zmud2_Cover)) {
  bareground[l] <- 100-sum(Zmud2_Cover[l])
}
Zmud2_Cover <- rbind(Zmud2_Cover, bareground)

rownames(Zmud2_Cover)[dim(Zmud2_Cover)[1]] <- "bareground"
```


```{r}

Zmud2_Dis<-vegdist(t(Zmud2_Cover), na.rm = TRUE, method = "bray") #Calculate Dissim between plots

Zmud2_CoverNMDS<-metaMDS(comm =t(Zmud2_Cover), distance = "bray"  ) #just doing a gross NMDS for visualization
ordiplot(Zmud2_CoverNMDS, type = "text")

Zmud2<- dist2list(Zmud2_Dis, tri = TRUE)
densityplot(Zmud2$value) 
densityplot(sqrt(Zmud2$value)) 

```
I'm not sure if adding in bareground was smart or not

## Environmental

```{r}
Zmud2_Env<-Zmud_Diss[c(1,7:9 )]
Zmud2_Env<- data.frame(Zmud2_Env[,-1], row.names=Zmud2_Env[,1])

Zmud2_Env_dissim <- vegdist(Zmud2_Env, method = "gower")
Zmud2_Env_dissim_long<- dist2list(Zmud2_Env_dissim, tri = TRUE)
Zmud2_EnvNMDS<-metaMDS(comm = Zmud2_Env, distance = "gower")
plot(Zmud2_EnvNMDS)
densityplot(Zmud2_Env_dissim_long$value)
densityplot(Zmud2_Env_dissim_long$value*.8)


```

## Aspect

Aspect has been such a fun puzzle to solve. Ultimately, I have decided on the two dissimilarity matrices route.

```{r}

Zmud_Aspect<- data.frame(Zmud_Diss$Aspect, row.names = Zmud_Diss[,1])


for(a in 1:nrow(CircleConverter)){
  Zmud_Aspect$Zmud_Diss.Aspect[which(Zmud_Aspect == CircleConverter$Aspect[a])] <- CircleConverter$Angle[a]

  
}

Zmud_Aspect$Zmud_Diss.Aspect <- as.numeric(Zmud_Aspect$Zmud_Diss.Aspect)
Zmud_Circle<- circular(Zmud_Aspect, units = "degrees", modulo = "2pi")
plot.circular(Zmud_Circle, stack = TRUE)
Zmud_CircleDiss<-dist.circular(Zmud_Circle, method = "geodesic")/pi #well this actually make the most sense
Zmud_CircleDissLong<- dist2list(Zmud_CircleDiss, tri = TRUE)
densityplot(Zmud_CircleDissLong$value)
summary(Zmud_CircleDissLong$value)
```

## Env Combined

```{r}
which(Zmud2_Env_dissim_long$col != Zmud_CircleDissLong$col)
which(Zmud2_Env_dissim_long$row != Zmud_CircleDissLong$row)

Zmud2$EnvDis<-(Zmud2_Env_dissim_long$value * 0.8) + (Zmud_CircleDissLong$value * 0.2)
which(Zmud2_Env_dissim_long$value == Zmud2$EnvDis)

densityplot(Zmud2$EnvDis)
summary(Zmud2$EnvDis)
summary(Zmud2$value)


ggplot(aes(x = EnvDis, y = value), data = Zmud2)+
  geom_point()+
  geom_smooth(method = "lm")

ggplot(aes(x = EnvDis, y = sqrt(value)), data = Zmud2)+
  geom_point()+
  geom_smooth(method = "lm")

summary(lm(value ~ EnvDis, data = Zmud2))
summary(lm(sqrt(value) ~ EnvDis, data = Zmud2))
```

## Actual Distances

```{r}


Zmud_difvalues<-setdiff(Zmud_RelCords$Plot, as.character(Zmud_Diss$Plot))



`%not_in%` <- purrr::negate(`%in%`)
Zmud2_RelCord<-Zmud_RelCords[Zmud_RelCords$Plot %not_in% Zmud_difvalues,]


Zmud2_RelCord <- data.frame(Zmud2_RelCord[,-1], row.names=Zmud2_RelCord[,1])
Zmud2_RelCord
Zmud2_X<-designdist(Zmud2_RelCord[1], method = "A-B", terms = "minimum")

Zmud2_Y<-designdist(Zmud2_RelCord[2], method = "A-B", terms = "minimum")
Zmud2_Z<-designdist(Zmud2_RelCord[3], method = "A-B", terms = "minimum")


Zmud2_Dist<-data.frame(dist2list(Zmud2_X, tri = TRUE)[1:2],Xdist = dist2list(Zmud2_X, tri = TRUE)$value, Ydist = dist2list(Zmud2_Y, tri = TRUE)$value, Zdist = dist2list(Zmud2_Z, tri = TRUE)$value)

Zmud2_Dist$TotalDist<-sqrt(Zmud2_Dist$Xdist^2 + Zmud2_Dist$Ydis^2 + Zmud2_Dist$Zdist^2)


densityplot(Zmud2_Dist$Xdist)
densityplot(Zmud2_Dist$Ydist)
densityplot(Zmud2_Dist$Zdist)
densityplot(Zmud2_Dist$TotalDist)

Zmud2<-data.frame(Zmud2_Dist[1:2], Dissim = Zmud2$value, Zmud2[4], Zmud2_Dist[3:6])


```

## Plots and models

```{r}

Zmud_EnvDis.mod <-lm(Dissim ~ EnvDis, data = Zmud2)
summary(Zmud_EnvDis.mod) # ***

ggplot(aes(y = Dissim, x = EnvDis ), data = Zmud2)+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Ten-Mile: Environmental Similarity between Points")


summary(lm(Dissim ~ abs(Xdist), data = Zmud2)) #***

ggplot(aes(y = Dissim, x = abs(Xdist) ), data = Zmud2)+
  geom_point()+
  geom_smooth(method = "lm")


summary(lm(Dissim ~ abs(Ydist), data = Zmud2)) #**

ggplot(aes(y = Dissim, x = abs(Ydist) ), data = Zmud2)+
  geom_point()+
  geom_smooth(method = "lm")

summary(lm(Dissim ~ (abs(Zdist)), data = Zmud2))

ggplot(aes(y = Dissim, x = sqrt(abs(Zdist)) ), data = Zmud2)+
  geom_point()


Zmud_TotalDist.mod<- lm(Dissim ~ TotalDist, data = Zmud2)
summary(Zmud_TotalDist.mod) # **

ggplot(aes(y = Dissim, x = TotalDist ), data = Zmud2)+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Ten-Mile: Distance between Points")


anova(Zmud_EnvDis.mod, Zmud_TotalDist.mod, test = "Chisq")

summary(lm(Dissim ~ EnvDis+TotalDist, data = Zmud2)) # **
summary(lm(Dissim ~ EnvDis*TotalDist, data = Zmud2)) # .

```



## Marginal Effects

```{r}
Zmud_Combined.mod <- lm(Dissim ~ TotalDist+EnvDis, data = Zmud2)
summary(Zmud_Combined.mod)
Zmud.pre<-predictions(Zmud_Combined.mod)
ggplot(aes(x = EnvDis, y = estimate), data = Zmud.pre)+
  geom_point()


Zmud.cmp<- avg_comparisons(Zmud_Combined.mod,  variables =  list(TotalDist = "minmax", EnvDis = "minmax"), conf_level = 0.90, type = "response")

Zmud.cmp

ggplot(Zmud.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", size = 1.5)+
  geom_errorbarh() + 
  geom_point(size = 2) +
  theme_classic()+
  labs(title = "Zmudowski")

plot_comparisons(Zmud_Combined.mod,
                 variables = "EnvDis",
                 condition = "TotalDist",
                 comparison = "ratio") 
```


# Guadalupe-Nipomo


## Cover

```{r}
GuNi2_Cover <- data.frame()

for (i in 1:length(GuNi_Names)) { #for-loop to combine focal and concentric quadrats 
  Temp_Cover<-GuNi_Cover %>% select(starts_with(paste0("GuNi.", GuNi_Names[i])))
  Temp_Concentric<-rowSums(Temp_Cover[2:5]/4) #CHECK THIS
  Temp_Cover2<- (Temp_Cover[1]*.25)+(Temp_Concentric*.75) #CHECK THIS
  if(i ==1) {
    GuNi2_Cover<- Temp_Cover2
  } else {
    GuNi2_Cover<-cbind(GuNi2_Cover, Temp_Cover2)
  }
}

mean(colSums(GuNi_Cover, na.rm = TRUE))
mean(colSums(GuNi2_Cover, na.rm = TRUE)) #I think these numbers lining up means it worked right
```


```{r}

#GuNi2_Cover <- GuNi2_Cover[,colSums(is.na(GuNi2_Cover))<nrow(GuNi2_Cover)]
GuNi2_Cover[is.na(GuNi2_Cover)] <- 0

colSums(GuNi2_Cover) > 100

bareground<- vector(mode="double", length=ncol(GuNi2_Cover))
names(bareground) <- "bareground"

for(l in 1:length(GuNi2_Cover)) {
  bareground[l] <- 100-sum(GuNi2_Cover[l])
}
GuNi2_Cover <- rbind(GuNi2_Cover, bareground)

rownames(GuNi2_Cover)[dim(GuNi2_Cover)[1]] <- "bareground"
```


```{r}

GuNi2_Dis<-vegdist(t(GuNi2_Cover), na.rm = TRUE, method = "bray") #Calculate Dissim between plots

GuNi2_CoverNMDS<-metaMDS(comm =t(GuNi2_Cover), distance = "bray"  ) #just doing a gross NMDS for visualization
ordiplot(GuNi2_CoverNMDS, type = "text")

GuNi2<- dist2list(GuNi2_Dis, tri = TRUE)
densityplot(GuNi2$value) 
densityplot(sqrt(GuNi2$value)) 

```
I'm not sure if adding in bareground was smart or not

## Environmental

```{r}
GuNi2_Env<-GuNi_Diss[c(1,7:9 , 14:16)]
GuNi2_Env<- data.frame(GuNi2_Env[,-1], row.names=GuNi2_Env[,1])

GuNi2_Env_dissim <- vegdist(GuNi2_Env, method = "gower", na.rm = TRUE)
GuNi2_Env_dissim_long<- dist2list(GuNi2_Env_dissim, tri = TRUE)
GuNi2_EnvNMDS<-metaMDS(comm = GuNi2_Env, distance = "gower", na.rm = TRUE)
plot(GuNi2_EnvNMDS)
densityplot(GuNi2_Env_dissim_long$value)
densityplot(GuNi2_Env_dissim_long$value*.8)


```

## Aspect

Aspect has been such a fun puzzle to solve. Ultimately, I have decided on the two dissimilarity matrices route.

```{r}

GuNi_Aspect<- data.frame(GuNi_Diss$Aspect, row.names = GuNi_Diss[,1])


for(a in 1:nrow(CircleConverter)){
  GuNi_Aspect$GuNi_Diss.Aspect[which(GuNi_Aspect == CircleConverter$Aspect[a])] <- CircleConverter$Angle[a]

  
}

GuNi_Aspect$GuNi_Diss.Aspect <- as.numeric(GuNi_Aspect$GuNi_Diss.Aspect)
GuNi_Circle<- circular(GuNi_Aspect, units = "degrees", modulo = "2pi")
plot.circular(GuNi_Circle, stack = TRUE)
GuNi_CircleDiss<-dist.circular(GuNi_Circle, method = "geodesic")/pi #well this actually make the most sense
GuNi_CircleDissLong<- dist2list(GuNi_CircleDiss, tri = TRUE)
densityplot(GuNi_CircleDissLong$value)
summary(GuNi_CircleDissLong$value)
```

## Env Combined

```{r}
which(GuNi2_Env_dissim_long$col != GuNi_CircleDissLong$col)
which(GuNi2_Env_dissim_long$row != GuNi_CircleDissLong$row)

GuNi2$EnvDis<-(GuNi2_Env_dissim_long$value * 0.8) + (GuNi_CircleDissLong$value * 0.2)
which(GuNi2_Env_dissim_long$value == GuNi2$EnvDis)

densityplot(GuNi2$EnvDis)
summary(GuNi2$EnvDis)
summary(GuNi2$value)


ggplot(aes(x = EnvDis, y = value), data = GuNi2)+
  geom_point()+
  geom_smooth(method = "lm")

ggplot(aes(x = EnvDis, y = sqrt(value)), data = GuNi2)+
  geom_point()+
  geom_smooth(method = "lm")

summary(lm(value ~ EnvDis, data = GuNi2))
summary(lm(sqrt(value) ~ EnvDis, data = GuNi2))
```

## Actual Distances

```{r}


GuNi_difvalues<-setdiff(GuNi_RelCords$Plot, as.character(GuNi_Diss$Plot))



`%not_in%` <- purrr::negate(`%in%`)
GuNi2_RelCord<-GuNi_RelCords[GuNi_RelCords$Plot %not_in% GuNi_difvalues,]


GuNi2_RelCord <- data.frame(GuNi2_RelCord[,-1], row.names=GuNi2_RelCord[,1])
GuNi2_RelCord
GuNi2_X<-designdist(GuNi2_RelCord[1], method = "A-B", terms = "minimum")

GuNi2_Y<-designdist(GuNi2_RelCord[2], method = "A-B", terms = "minimum")
GuNi2_Z<-designdist(GuNi2_RelCord[3], method = "A-B", terms = "minimum")


GuNi2_Dist<-data.frame(dist2list(GuNi2_X, tri = TRUE)[1:2],Xdist = dist2list(GuNi2_X, tri = TRUE)$value, Ydist = dist2list(GuNi2_Y, tri = TRUE)$value, Zdist = dist2list(GuNi2_Z, tri = TRUE)$value)

GuNi2_Dist$TotalDist<-sqrt(GuNi2_Dist$Xdist^2 + GuNi2_Dist$Ydis^2 + GuNi2_Dist$Zdist^2)


densityplot(GuNi2_Dist$Xdist)
densityplot(GuNi2_Dist$Ydist)
densityplot(GuNi2_Dist$Zdist)
densityplot(GuNi2_Dist$TotalDist)

GuNi2<-data.frame(GuNi2_Dist[1:2], Dissim = GuNi2$value, GuNi2[4], GuNi2_Dist[3:6])


```

## Plots and models

```{r}

GuNi_EnvDis.mod <-lm(Dissim ~ EnvDis, data = GuNi2)
summary(GuNi_EnvDis.mod) # ***

ggplot(aes(y = Dissim, x = EnvDis ), data = GuNi2)+
  geom_point()+
  geom_smooth(method = "lm")+
  labs(title = "Ten-Mile: Environmental Similarity between Points")


summary(lm(Dissim ~ abs(Xdist), data = GuNi2))

ggplot(aes(y = Dissim, x = abs(Xdist) ), data = GuNi2)+
  geom_point()+
  geom_smooth(method = "lm")


summary(lm(Dissim ~ abs(Ydist), data = GuNi2))

ggplot(aes(y = Dissim, x = abs(Ydist) ), data = GuNi2)+
  geom_point()

summary(lm(Dissim ~ (abs(Zdist)), data = GuNi2)) # **

ggplot(aes(y = Dissim, x = (abs(Zdist)) ), data = GuNi2)+
  geom_point()


GuNi_TotalDist.mod<- lm(Dissim ~ TotalDist, data = GuNi2)
summary(GuNi_TotalDist.mod) # 

ggplot(aes(y = Dissim, x = TotalDist ), data = GuNi2)+
  geom_point()+
  labs(title = "Ten-Mile: Distance between Points")


anova(GuNi_EnvDis.mod, GuNi_TotalDist.mod, test = "Chisq")

summary(lm(Dissim ~ EnvDis+TotalDist, data = GuNi2)) # **
summary(lm(Dissim ~ EnvDis*TotalDist, data = GuNi2)) # .

```



## Marginal Effects

```{r}
GuNi_Combined.mod <- lm(Dissim ~ TotalDist+EnvDis, data = GuNi2)
summary(GuNi_Combined.mod)
GuNi.pre<-predictions(GuNi_Combined.mod)
ggplot(aes(x = EnvDis, y = estimate), data = GuNi.pre)+
  geom_point()


GuNi.cmp<- avg_comparisons(GuNi_Combined.mod,  variables =  list(TotalDist = "minmax", EnvDis = "minmax"), conf_level = 0.90, type = "response")

GuNi.cmp

ggplot(GuNi.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", size = 1.5)+
  geom_errorbarh() + 
  geom_point(size = 2) +
  theme_classic()+
  labs(title = "Guadalupe-Nipomo")

ggplot(aes(x = EnvDis, y = Dissim), data = GuNi2)+
  geom_point()+
  geom_smooth(method = "lm")

```


# **Combined Analysis**

### Data management

```{r}

TeMi2$Site<- rep("TeMi", dim(TeMi2)[1])
Zmud2$Site<- rep("Zmud", dim(Zmud2)[1])
GuNi2$Site<- rep("GuNi", dim(GuNi2)[1])

Combo2<-rbind(TeMi2, Zmud2, GuNi2)
Combo2$Site <- factor(Combo2$Site, levels=c('TeMi', 'Zmud', 'GuNi'))

densityplot(~ Dissim, data = Combo2)
densityplot(~ Dissim, group = Site, data = Combo2)
summary(lm(Dissim ~ Site, data = Combo2)) # yeah it is sig, but that is likely just a product of the large sample size rather than a meaningful difference
ggplot(aes(x = Site, y = Dissim, color = Site), data = Combo2)+
  geom_jitter(width = 0.1, height = 0)+
  geom_violin(aes(fill = Site), alpha = 0.5)+
  stat_summary_bin(fun = "mean", geom = "point", orientation = 'x', size = 5, color = "black")


ggplot(aes(x = EnvDis, y = Dissim), data = Combo2)+
  geom_point(aes(color = Site))+
  geom_smooth(method = "lm", color = "black")+
  geom_smooth(aes(color = Site),method = "lm")



```

### Negligible effects
```{r}
Mixed2<- lmer(Dissim ~ EnvDis+TotalDist + (1|Site), data = Combo2)
summary(Mixed2)
car::Anova(Mixed2)

Mixed2.Inter<- lmer(Dissim ~ EnvDis*TotalDist + (1|Site), data = Combo2)

Mixed2.null<- lmer(Dissim ~ 1 + (1|Site), data = Combo2)
anova(Mixed2, Mixed2.null)
anova(Mixed2, Mixed2.Inter)
Mixed2.cmp<- avg_comparisons(Mixed2, variables = list(EnvDis = "minmax", TotalDist = "minmax"),   conf_level = 0.90, type = "response")


as.data.frame(Mixed2.cmp)

Bet.Negl<- ggplot(Mixed2.cmp, aes(x = estimate,
                 xmin = conf.low,
                 xmax = conf.high, 
                 y = term,
                 group = contrast)) +
     annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, alpha = .2)+
  geom_vline(xintercept = c(-0.1, .1), linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "solid", linewidth = 1.5)+
  geom_errorbarh(aes(xmin = conf.low,  xmax = conf.high,),position = position_dodgev(height = 1.5)) + 
  geom_point(size = 4, position = position_dodgev(height = 1.5)) +
  theme_classic()+
  labs(x = "Estimated Change in Community Dissimilarity", y = "Term")+
  theme(text = element_text(family = "serif", size = 18), axis.text.y = element_text( size=12, angle=45))+ 
  scale_y_discrete(labels=c("TotalDist" = "Distance Between Plots", 
                            "EnvDis" = "Environmental Dissimilarity"))

Bet.Negl
ggsave("Between-Marginal.png", dpi = 600, height = 7, width = 10)


```

```{r}
plot_predictions(Mixed2, condition = "EnvDis")+
  geom_point(data = Combo2,alpha = 0.5, shape = 1, aes(x = EnvDis, y = Dissim, color = Site))+
  labs(x = "Environmental Dissimilarity", y =  "Community Dissimilarity")+
  scale_color_manual(labels = c("Guadalupe-Nipomo", "Ten Mile", "Zmudowski"), values = c("#d55e00", "#cc79a7", "#009e73"))+
  theme(text = element_text(family = "serif", size = 18))



library(ggside)
pp_between_EnvDis<-plot_predictions(Mixed2, condition = "EnvDis", draw = FALSE)


Combo3 <- Combo2 #this is dangerous and bad
Combo3$estimate <- Combo3$Dissim


Bet.EnvDis <- ggplot(aes(x = EnvDis, y = estimate), data = pp_between_EnvDis)+
  geom_point(data = Combo2,alpha = 0.25, aes(x = EnvDis, y = Dissim, shape = Site))+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.5, color = "transparent", fill = "grey")+
  geom_line(size = .5)+
  labs(x = "Environmental Dissimilarity", y =  "Community Dissimilarity")+
  scale_color_manual(labels = c("Ten Mile", "Zmudowski", "Guadalupe-Nipomo"), values = c("#d55e00", "#cc79a7", "#009e73"), name = "Site")+
  scale_shape_manual(labels = c("Ten Mile", "Zmudowski", "Guadalupe-Nipomo"), values = c(1,2,0), name = "Site")+
  theme_classic()+
  theme(text = element_text(family = "serif", size = 18),legend.position='bottom', legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid'))+
    geom_xsidedensity(aes(y= after_stat(density),color = Site, fill = Site),
    alpha    = 0.25,
    size     = 1,
    position = "identity",
    data = Combo3
  ) +
    geom_ysidedensity(aes(x= after_stat(density),color = Site, fill = Site),
    alpha    = 0.25,
    size     = 1,
    position = "identity",
    data = Combo3
  ) +
  scale_fill_manual(labels = c("Ten Mile", "Zmudowski", "Guadalupe-Nipomo"), values = c("#d55e00", "#cc79a7", "#009e73"), name = "Site")+
  theme_ggside_void()

Bet.EnvDis


ggsave("Between-EnvDis.png", dpi = 600, height = 7, width = 10)


pp_between_TotalDist<-plot_predictions(Mixed2, condition = "TotalDist", draw = FALSE)

Bet.TotalDist<- ggplot(aes(x = TotalDist, y = estimate), data = pp_between_TotalDist)+
  geom_point(data = Combo2,alpha = 0.25, aes(x = TotalDist, y = Dissim, shape = Site))+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.5, color = "transparent", fill = "grey")+
  geom_line(size = .5)+
  labs(x = "Distance Between Plots (m)", y =  "Community Dissimilarity")+     
  scale_fill_manual(labels = c("Ten Mile", "Zmudowski", "Guadalupe-Nipomo"), values = c("#d55e00", "#cc79a7", "#009e73"), name = "Site")+
  scale_color_manual(labels = c("Ten Mile", "Zmudowski", "Guadalupe-Nipomo"), values = c("#d55e00", "#cc79a7", "#009e73"), name = "Site")+
  scale_shape_manual(labels = c("Ten Mile", "Zmudowski", "Guadalupe-Nipomo"), values = c(1,2,0), name = "Site")+
  theme_classic()+
  theme(text = element_text(family = "serif", size = 18),legend.position='bottom', legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid'))+
    geom_xsidedensity(aes(y= after_stat(density),color = Site, fill = Site),
    alpha    = 0.25,
    size     = 1,
    position = "identity",
    data = Combo3
  ) +
    geom_ysidedensity(aes(x= after_stat(density),color = Site, fill = Site),
    alpha    = 0.25,
    size     = 1,
    position = "identity",
    data = Combo3
  ) +
  theme_ggside_void()


Bet.TotalDist


ggsave("Between-EnvDis.png", dpi = 600, height = 7, width = 10)

```

# Figure Construction

```{r}

Bet.TotalDist2<- Bet.TotalDist + theme(legend.position='none')
ggdraw() +
  draw_plot(Bet.Negl, x = 0, y = .666, width = 1, height = .333) +
  draw_plot(Bet.TotalDist2, x = 0, y = 0.3333, width = 1, height = .333) +
  draw_plot(Bet.EnvDis, x = 0, y = 0, width = 1, height = 0.3333) +
  draw_plot_label(label = c("A", "B", "C" ),family = "serif", size = 20, color = "black", x = c(0, 0, 0), y = c(.9999, .69, 0.355 ))+
  draw_image("BetweenPlotLogo.png", x = 0.79, y = 0.83, width = .175, height = 0.175) +
  theme(plot.background = element_rect(fill = "white", color = "transparent"))

ggsave("BetweenGraph.png", dpi = 600, height = 12, width = 9)


```

# Tables

```{r}
models.plots2<- list( 
  "Reduced Model" = Mixed2,
  "Full Model" = Mixed2.Inter,
  "Null Model" = Mixed2.null
   )

fun <- function(model) data.frame("corrected" = round(cAIC4::cAIC(model)$caic, 2))


modelsummary(models.plots2,
  fmt = 3,
  estimate  = "{estimate} [{conf.low}, {conf.high}]",
  statistic = NULL, 
  conf_level = .90,
  gof_function = fun, 
  gof_omit = 'AIC|BIC|ICC',
  output = "BetweenPlotsTable.docx")




```


```{r}
save.image(file="DuneComm4.RData") 
```

