---
title: "Categorical Aspect and Dissimilarity Exploration"
author: "Ben J. Rivera"
date: '2023-01-31'
output: html_document
---

Are you like me where you come across incredibly specific problems with ecological data that nobody seems to be talking about despite it seeming like it should be common? If you are, then stick around because you might like this! 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(vegan)
library(circular)
library(climaemet)

load(file = ".RData") #this allows me to use the global data. So I can make dataframes elsewhere and then use them here.

```


```{r}

plotsetup<- data.frame(direction= c("N", "NE", "E", "SE", "S", "SW","W","NW"), 
                       angle = seq(0,315, 45) )

ggplot(plotsetup)+
  coord_polar()+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  annotate("text",label = plotsetup$direction,x = plotsetup$angle, y = 1)




```

# Four Column Solution
```{r}

FourCol <- data.frame(row.names = c("N", "NE", "E", "SE", "S", "SW","W","NW"), 
                      N = c(1,1,0,0,0,0,0,1), 
                      S = c(0,0,0,1,1,1,0,0), 
                      W = c(0,0,0,0,0,1,1,1), 
                      E = c(0,1,1,1,0,0,0,0))

FourDis<- vegdist(FourCol, method = "gower", upper = TRUE) # 
FourDis
dissimsums1<-colSums(as.matrix(FourDis))
dissimsums1


FourDisNorth<-cbind(plotsetup, dissim = as.matrix(FourDis)[,1])

FourDisNE<- cbind(plotsetup, dissim = as.matrix(FourDis)[,2])


ggplot(data = FourDisNorth, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Four Column Solution", subtitle = "North's Comparisons", y = "Calculated Dissimilarity")+
  scale_fill_discrete(guide = "none")

ggplot(data = FourDisNE, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Four Column Solution", subtitle = "NorthEast's Comparisons", y = "Calculated Dissimilarity")+
  scale_fill_discrete(guide = "none")
```

```{r}

FourColHalf<- data.frame(row.names = c("N", "NE", "E", "SE", "S", "SW","W","NW"), 
                      N = c(1,0.5,0,0,0,0,0,0.5), 
                      S = c(0,0,0,0.5,1,0.5,0,0), 
                      W = c(0,0,0,0,0,0.5,1,0.5), 
                      E = c(0,0.5,1,0.5,0,0,0,0))

FourColHalf
rowSums(FourColHalf)
FourDisHalf<- vegdist(FourColHalf, method = "gower", upper = TRUE) # bray-curtis dissim
FourDisHalf
as.matrix(FourDisHalf)
colSums(as.matrix(FourDisHalf))

FourHalfNorth<-cbind(plotsetup, dissim = as.matrix(FourDisHalf)[,1])

FourHalfNE<- cbind(plotsetup, dissim = as.matrix(FourDisHalf)[,2])

cor(FourDis, FourDisHalf)


ggplot(data = FourHalfNorth, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Four Column with .5's Solution", subtitle = "North's Comparisons")+
  scale_fill_discrete(guide = "none")

ggplot(data = FourHalfNE, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Four Column with .5's Solution", subtitle = "NorthEast's Comparisons", y = "Calculated Dissimilarity")+
  scale_fill_discrete(guide = "none")


```

```{r}
FourColWeird<- data.frame(row.names = c("N", "NW","W","SW", "S", "SE","E","NE"), 
                      N_S = c(1,0.75,0.5,0.25,0,0.25,0.5,0.75), 
                      NW_SE = c(0.75,1,0.75,0.5,0.25,0,0.25,0.5), 
                      E_W = c(0.5,0.25,0,0.25,0.5,0.75,1,0.75), 
                      NE_SW = c(0.75,0.5,0.25,0,0.25,0.5,0.75,1))

FourColWeird
rowSums(FourColWeird)

FourDisWeird<- vegdist(FourColWeird, method = 'gower', upper = TRUE)


FourDisWeird #HOLY SHIT I THINK THIS IS THE ONE THAT ACTUALLY WORKS WTF
colSums(as.matrix(FourDisWeird))

FourWeirdNorth<-cbind(plotsetup, dissim = as.matrix(FourDisWeird)[,1])

FourWeirdNE<- cbind(plotsetup, dissim = as.matrix(FourDisWeird)[,2])

cor(FourDis, FourDisHalf)


ggplot(data = FourWeirdNorth, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Four Column with .25's Solution", subtitle = "North's Comparisons")+
  scale_fill_discrete(guide = "none")

ggplot(data = FourWeirdNE, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Four Column with .25's Solution", subtitle = "NorthEast's Comparisons", y = "Calculated Dissimilarity")+
  scale_fill_discrete(guide = "none")

```

Hmmm, N and NW are .333 dissimilar which is okay. SE and NE and .5 dissimilar which I don't quite understand why that isn't also .333. Hmmmm

#Two Column solution


SOMETHING IS WRONG HERE
```{r}
TwoCol <- data.frame(row.names = c("N", "NW","W","SW", "S", "SE","E","NE"), 
                     NS = c(1,1,0.5,0,0,0,0.5,1),
                     EW = c(0.5,0,0,0,0.5,1,1,1))
TwoCol

TwoDis<- vegdist(TwoCol, method = "gower", upper = TRUE) # bray-curtis dissim
round(TwoDis, 2)


dist(FourCol, method = "binary")
(TwoDis == FourDis)

TwoHalfNorth<-cbind(plotsetup, dissim = as.matrix(TwoDis)[,1])

TwoHalfNE<- cbind(plotsetup, dissim = as.matrix(TwoDis)[,2])



ggplot(data = TwoHalfNorth, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Two Column with .5's Solution", subtitle = "North's Comparisons")+
  scale_fill_discrete(guide = "none")

ggplot(data = TwoHalfNE, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Two Column with .5's Solution", subtitle = "NorthEast's Comparisons", y = "Calculated Dissimilarity")+
  scale_fill_discrete(guide = "none")

```

```{r}
TwoColNeg<- data.frame(row.names = c("N", "NW","W","SW", "S", "SE","E","NE"), 
                     NS = c(1,1,0,-1,-1,-1,0,1),
                     EW = c(0,-1,-1,-1,0,1,1,1))

TwoColNeg
TwoDisNeg<- vegdist(TwoColNeg, method = "gower", upper = TRUE) # bray-curtis dissim
TwoDisNeg

(TwoDis == TwoDisNeg)

TwoNegNorth<-cbind(plotsetup, dissim = as.matrix(TwoDisNeg)[,1])

TwoNegNE<- cbind(plotsetup, dissim = as.matrix(TwoDisNeg)[,2])

ggplot(data = TwoNegNorth, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Two Column with Negatives Solution", subtitle = "North's Comparisons")+
  scale_fill_discrete(guide = "none")

ggplot(data = TwoNegNE, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Two Column with Negatives Solution", subtitle = "NorthEast's Comparisons", y = "Calculated Dissimilarity")+
  scale_fill_discrete(guide = "none")

```
# oops 3 column

```{r}
ThreeCol<- data.frame(row.names = c("N", "NW","W","SW", "S", "SE","E","NE"), 
                      N_S = c(1,0.75,0.5,0.25,0,0.25,0.5,0.75), 
                      NW_SE = c(0.75,1,0.75,0.5,0.25,0,0.25,0.5), 
                      E_W = c(0.5,0.25,0,0.25,0.5,0.75,1,0.75) 
                     )

ThreeCol
rowSums(FourColWeird)

ThreeDis<- vegdist(ThreeCol, method = 'gower')


ThreeDis #not even close


ThreeColNorth<-cbind(plotsetup, dissim = as.matrix(ThreeDis)[,1])

ThreeColNE<- cbind(plotsetup, dissim = as.matrix(ThreeDis)[,2])

ggplot(data = ThreeColNorth, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Three Column Solution", subtitle = "North's Comparisons")+
  scale_fill_discrete(guide = "none")

ggplot(data = ThreeColNE, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Three Column Solution", subtitle = "NorthEast's Comparisons", y = "Calculated Dissimilarity")+
  scale_fill_discrete(guide = "none")

```


# circular

```{r}

Circle <- data.frame(row.names = c("N", "NW","W","SW", "S", "SE","E","NE"), 
                     Angle = c(0,45,90,135,180,225,270,315))

Circle
dist.circular(Circle)
Circ<-circular(Circle, rotation = "clock")
dist.circular(Circ)


Cic<-circular(Circle, units = "degrees", modulo = "2pi", rotation = "clock", template = "geographics")
Cic

plot.circular(Cic)

CircleDiss<-dist.circular(Cic, method = "geodesic", upper = TRUE)/pi #well this actually make the most sense
CircleDiss

CircleNorth<-cbind(plotsetup, dissim = as.matrix(CircleDiss)[,1])

CircleNE<- cbind(plotsetup, dissim = as.matrix(CircleDiss)[,2])



ggplot(data = CircleNorth, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Circle Solution", subtitle = "North's Comparisons")+
  scale_fill_discrete(guide = "none")

ggplot(data = CircleNE, aes(x = angle, y = dissim, color = dissim, fill = direction))+
  coord_polar(clip = "off")+
  scale_x_continuous(limits = c(0,360))+
  scale_y_continuous(limits = c(0,1))+
  geom_label(aes(label = direction, x = angle, y =1), color = "black")+
  geom_segment(aes(y = 0, yend = dissim), size = 1.5)+
  geom_point(size= 5, shape = 21)+ 
  scale_color_gradient2( high="red", 
                        limits = c(0, 1), name = "Dissimilarity")+
  labs(title = "Circle Solution", subtitle = "NorthEast's Comparisons", y = "Calculated Dissimilarity")+
  scale_fill_discrete(guide = "none")


```



