---
title: "Test"
author: "Ben J. Rivera"
date: '2023-01-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
load(file = ".RData") #this allows me to use the global data. So I can make dataframes elsewhere and then use them here.
head(GuNi_Locations)
```

# PCA reconscturction idea

stole this from here. Might be interesting to reconstruct predictions from some of these models and I found this. Basically, I would use PCA to reduce the number of axes then put those axes in a GLMM (needed because of low sample size at high spatial scales). After that, I can create predictions. But those predictions would be in made up PCA units. Turns out that you kinda sorta can back transform PCA numbers into real numbers a litte. Idk. We'll see.

here is an example I stole from here: https://stats.stackexchange.com/questions/229092/how-to-reverse-pca-and-reconstruct-original-variables-from-several-principal-com 


```{r}
X = iris[,1:4]

mu = colMeans(X)


Xpca = prcomp(X)

nComp = 2

plot(Xpca$x[,1:nComp])
Xhat = Xpca$x[,1:nComp] %*% t(Xpca$rotation[,1:nComp])
Xhat = scale(Xhat, center = -mu, scale = FALSE)

iris[1,];Xhat[1,] #first is real, second is reconstructed
```

